\documentclass{article}
\usepackage{iclr2026_conference,times}
\input{math_commands.tex}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{graphicx}
\usepackage{array}
\usepackage{xcolor}
\usepackage{enumitem}
\usepackage[breaklinks=true,colorlinks=false]{hyperref}
\hypersetup{
  pdfborder={0 0 1},
  citebordercolor={0 1 0},
  linkbordercolor={1 0 0},
  urlbordercolor={0 1 0}
}

\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

\newcommand{\method}{Branch-Counterfactual MPC}
\newcommand{\methodv}{Branch-Counterfactual MPC v5}
\newcommand{\latent}{z}
\newcommand{\obs}{o}
\newcommand{\act}{a}
\newcommand{\state}{x}
\newcommand{\loss}{\ell}
\newcommand{\belief}{b}
\newcommand{\horizon}{H}
\newcommand{\dataset}{\mathcal{D}}
\newcommand{\branchset}{\mathcal{Z}}
\newcommand{\actionset}{\mathcal{A}}
\newcommand{\expect}{\mathbb{E}}

\title{WAM Contact Counterfactuals: Branch-Structured Planning for Latent Contact Dynamics}
\author{Anonymous Authors}

\begin{document}
\maketitle

\begin{abstract}
Contact-rich manipulation is not only a prediction problem.  A single executed push hides many physically plausible alternatives: the object may be heavier, the contact may slip earlier, or an apparently safe action may enter a different contact mode.  This paper studies whether a world-action model for contact-rich manipulation should keep such alternatives explicit until physical observations collapse them.  We rebuild WAM Contact Counterfactuals as a real MuJoCo contact-pushing study with latent mass/friction branches, strong MPC baselines, posterior diagnostics, adversarial stress splits, and a frozen evaluation protocol.  The proposed \methodv{} chooses diagnostic actions by bounded expected information gain, updates a posterior over contact branches, and falls back to conservative planning when the latent mode remains underidentified.  The paper is intentionally evidence-gated: branch structure is a contribution only if it beats ensemble, adaptive-identification, and robust MPC controls under paired hidden-physics tasks.  If the final frozen evidence does not clear that bar, the correct outcome is \textbf{STRONG\_REVISE}, not a polished but fragile submission claim.
\end{abstract}

\section{Introduction}

Robot manipulation papers often report the trajectory that happened.  Contact-rich manipulation also depends on trajectories that could have happened but did not: a pusher might have stuck instead of slipped, a small change in friction might have made the same action underdrive the puck, or a heavier object might have moved just enough during a diagnostic contact to mislead a nominal planner.  These unrealized alternatives are not decorative uncertainty.  They are action-relevant latent futures.

The central question in this paper is deliberately narrow:
\begin{quote}
\emph{When a robot has a small action budget under hidden contact dynamics, does preserving branch-structured counterfactual futures improve planning beyond averaging or robustifying over the same physical models?}
\end{quote}
This question matters because the surrounding literature is crowded.  Contact-rich manipulation has strong work on diffusion policies, tactile and force-aware world models, visuo-tactile demonstrations, residual correction, and planning-guided policy learning \citep{diffusionVLA2026,omniVTA2026,dreamTacVLA2025,forceVLA2025,dexvitac2026,planningGuidedDiffusion2024,tactileSkins2024}.  A paper cannot be novel merely because it adds uncertainty to a planner.  It must show that the uncertainty is represented in a way that changes decisions and survives strong controls.

We therefore treat this as a hostile-review exercise rather than a showcase.  Earlier versions of this project were too small and too synthetic.  The v4 rebuild used real MuJoCo contact dynamics and showed an uncomfortable result: branch MPC beat random and nominal MPC but did not clearly separate from ensemble mean MPC or robust worst-case MPC.  The present version expands the method, theory, diagnostics, and final protocol.  It also keeps the acceptance gate explicit: if the expanded method still ties strong baselines, the contribution is not ICLR-main-ready.

\paragraph{Contributions.}
This paper contributes:
\begin{enumerate}[leftmargin=1.4em]
  \item A formal latent-contact planning setup in which a finite branch library represents physically plausible counterfactual contact futures.
  \item A branch-structured MPC method with bounded information-gain diagnostics, posterior reweighting, and entropy-triggered conservative fallback.
  \item A negative identifiability analysis explaining when branch structure cannot beat ensemble or robust controls because contact observations do not distinguish latent modes within the action budget.
  \item A CPU-only MuJoCo benchmark with paired hidden-physics tasks, strong MPC baselines, adversarial shifts, ablations, posterior diagnostics, bootstrap intervals, paired sign-flip tests, and oracle regret.
  \item A submission-readiness decision rule that reports all frozen results honestly and rejects the paper if the mechanism does not beat strong controls.
\end{enumerate}

\paragraph{What this paper does not claim.}
The method is not a learned visuo-tactile policy, not a hardware result, and not a universal contact manipulation benchmark.  It is a controlled mechanism test.  That choice is intentional: before scaling to richer perception and hardware, the mechanism must survive the simpler question of whether branch-explicit latent contact planning beats strong planners with access to the same branch library.

\section{Related Work and Hostile Prior-Work Boundary}

\paragraph{World-action models and contact-rich manipulation.}
Recent work explores world models, force-aware policies, tactile prediction, and multimodal representations for contact-rich manipulation \citep{omniVTA2026,dreamTacVLA2025,forceVLA2025,tacforesight2026,fawam2026}.  These systems often improve contact performance by adding sensory channels, larger model classes, or richer demonstrations.  Our work asks a different mechanistic question: if a planner already has a set of plausible contact dynamics, should it collapse them to an average prediction, guard against the worst branch, identify a maximum-likelihood branch, or preserve branch structure as an object of planning?

\paragraph{Diffusion, VLA, and data augmentation.}
Diffusion policies and vision-language-action systems are powerful baselines for many manipulation regimes \citep{planningGuidedDiffusion2024,diffusionVLA2026,forceVLA2025}.  Data augmentation and demonstration scaling can cover more outcomes, but they do not by themselves specify how a low-budget controller should trade immediate progress against diagnostic contact.  The present method can be read as a small, explicit planning module that tests whether counterfactual contact alternatives should be kept separate until an observation discriminates them.

\paragraph{Tactile and force sensing.}
Tactile and force-aware methods reduce ambiguity by observing contact more directly \citep{dexvitac2026,tactileSkins2024,tacforesight2026}.  We view such sensing as complementary.  In fact, our identifiability analysis predicts that branch-structured planning becomes more valuable when observations separate latent contact modes and less valuable when they do not.  A reviewer could fairly ask for tactile hardware validation; this paper does not hide that limitation.

\paragraph{Robust and belief-space planning.}
Robust MPC protects against poor outcomes by optimizing worst-case or tail-risk objectives \citep{mayne2000mpc}.  POMDP and belief-space planning maintain distributions over latent state \citep{kaelbling1998planning}.  Our contribution is not the existence of belief updates.  The contribution, if supported, is the contact-specific decision structure: a small set of branch rollouts, a diagnostic action constrained by progress regret, posterior diagnostics tied to physical parameters, and an explicit comparison against ensemble, robust, and adaptive-identification MPC using identical MuJoCo tasks.

\paragraph{The hostile novelty boundary.}
The local hostile prior-work audit identifies the crowded territory as larger models, generic uncertainty, benchmark-only claims, and data augmentation.  The defensible boundary is narrower: branch-explicit latent contact alternatives must change closed-loop decisions in a way observed-only or collapsed-belief controls cannot match.  If final paired evidence does not show that, this paper remains a strong empirical scaffold rather than a main-conference submission.

\section{Latent Contact Planning Setup}

We model a contact-rich manipulation episode as a finite-horizon controlled process with hidden physical mode $\latent$.  In the MuJoCo benchmark, $\latent$ corresponds to object mass and contact friction, but the formalism also covers contact-mode support, slip parameters, and unobserved compliance.

\begin{definition}[Latent contact planning problem]
A latent contact planning problem is a tuple
\[
  \mathcal{P} = (\mathcal{X}, \mathcal{A}, \mathcal{Z}, T_{\latent}, O, \loss, \horizon, b_0),
\]
where $\mathcal{X}$ is the observed physical state space, $\mathcal{A}$ is an action space, $\mathcal{Z}$ is a latent branch set, $T_{\latent}(\state'|\state,\act)$ is the branch-conditioned transition model, $O(\obs|\state,\latent)$ is an observation model, $\loss(\state,\act,\latent)$ is a task loss, $\horizon$ is the action budget, and $b_0$ is a prior over branches.
\end{definition}

At time $t$, the planner observes $\obs_t$, maintains a belief $b_t(\latent)$, chooses action $\act_t$, observes $\obs_{t+1}$, and updates
\[
  b_{t+1}(\latent) =
  \frac{b_t(\latent) p(\obs_{t+1} \mid \obs_t,\act_t,\latent)}
  {\sum_{\latent' \in \branchset} b_t(\latent') p(\obs_{t+1} \mid \obs_t,\act_t,\latent')}.
\]
In our implementation, the likelihood is induced by branch rollout error:
\[
  p(\obs_{t+1} \mid \obs_t,\act_t,\latent)
  \propto
  \exp\left(-\frac{\|f_{\latent}(\obs_t,\act_t)-\obs_{t+1}\|_2^2}{2\sigma^2}\right),
\]
followed by a small uniform mixing term to avoid irreversible posterior collapse after noisy contacts.

\paragraph{Branch values.}
For a candidate action $\act$ and branch $\latent$, let
\[
  d_{\latent}(\act) =
  \left\| f_{\latent}(\obs_t,\act)-g \right\|_2
\]
be the predicted distance to target $g$.  The planners compared in this paper differ in how they aggregate $\{d_{\latent}(\act)\}_{\latent \in \branchset}$:
\begin{align}
  J_{\mathrm{mean}}(\act) &= \sum_{\latent} b_t(\latent)d_{\latent}(\act),\\
  J_{\mathrm{worst}}(\act) &= \max_{\latent} d_{\latent}(\act),\\
  J_{\mathrm{hybrid}}(\act) &= \alpha J_{\mathrm{mean}}(\act) + (1-\alpha)J_{\mathrm{worst}}(\act),\\
  J_{\mathrm{tail}}(\act) &= \expect_{\latent \sim b_t}\left[d_{\latent}(\act) \mid d_{\latent}(\act) \text{ in upper posterior tail}\right].
\end{align}
Nominal MPC uses a single fixed branch.  Ensemble MPC minimizes $J_{\mathrm{mean}}$ with a uniform belief.  Robust MPC minimizes $J_{\mathrm{worst}}$.  Adaptive-ID MPC updates the belief but plans under the maximum-posterior branch.  \methodv{} uses diagnostic actions and then plans with tail or hybrid risk depending on posterior entropy.

\section{Method}

\subsection{Branch-Counterfactual MPC v5}

\methodv{} has three components.  First, it keeps a small branch library of plausible physical dynamics.  Second, it chooses a bounded diagnostic action that is predicted to reduce branch entropy without sacrificing too much immediate progress.  Third, it updates branch weights after contact and falls back to conservative planning when observations do not identify the branch.

\paragraph{Bounded diagnostic action.}
For each candidate action $\act$, the planner computes branch predictions
\[
  y_{\latent}(\act) = f_{\latent}(\obs_t,\act).
\]
The expected posterior entropy after $\act$ is approximated by treating each branch prediction $y_{\latent}$ as a possible observation:
\[
  \widehat{H}_{t+1}(\act)
  =
  \sum_{\latent \in \branchset} b_t(\latent)
  H\left(
  \mathrm{BayesUpdate}(b_t,\act,y_{\latent}(\act))
  \right).
\]
The information gain is
\[
  \mathrm{IG}(\act) = H(b_t)-\widehat{H}_{t+1}(\act).
\]
We do not allow arbitrary information-seeking.  Let $J_{\min}=\min_{\act'}J_{\mathrm{mean}}(\act')$.  The diagnostic safe set is
\[
  \mathcal{S}_{\epsilon}
  =
  \{\act \in \actionset: J_{\mathrm{mean}}(\act) \le J_{\min}+\epsilon\}.
\]
\methodv{} selects
\[
  \act^{\mathrm{diag}}
  =
  \arg\max_{\act \in \mathcal{S}_{\epsilon}}
  \left[
    \lambda_{\mathrm{ig}}\mathrm{IG}(\act)
    +\lambda_{\mathrm{prog}}\mathrm{Prog}(\act)
    -\lambda_{\mathrm{risk}}J_{\mathrm{mean}}(\act)
  \right].
\]
This construction is intentionally conservative.  A diagnostic action must be useful for identification and still remain near the best predicted progress action.

\paragraph{Entropy-triggered fallback.}
After each contact, the method records posterior entropy.  If entropy remains high after a diagnostic contact, the method treats the problem as underidentified and switches to a robust-mean hybrid objective.  This is not hidden tuning: fallback activations are reported as a metric.  A method that frequently falls back is admitting that the branch evidence did not collapse quickly enough.

\paragraph{Algorithm.}
For each episode:
\begin{enumerate}[leftmargin=1.4em]
  \item Initialize a uniform belief over the branch library.
  \item On the first push, choose the bounded information-gain diagnostic action.
  \item Execute the MuJoCo push under the true hidden mass/friction and any split-specific noise.
  \item Update the branch posterior from observed displacement.
  \item For remaining pushes, plan with tail risk if entropy is low and robust-mean hybrid risk if entropy is high.
  \item Report task metrics and posterior diagnostics.
\end{enumerate}

\subsection{Baselines}

The final protocol compares against:
\begin{itemize}[leftmargin=1.4em]
  \item random push, a weak sanity baseline;
  \item nominal single-branch MPC;
  \item ensemble mean MPC, which averages over the same branch library;
  \item domain-randomized MPC, which replans under sampled branch subsets;
  \item adaptive-ID MPC, which updates the posterior and then plans under the maximum-posterior branch;
  \item robust worst-case MPC;
  \item robust-mean hybrid MPC;
  \item the legacy v4 branch MPC;
  \item oracle hidden-parameter MPC, an upper bound that knows the true physical parameters.
\end{itemize}
These baselines are intentionally uncomfortable.  If the proposed method only beats nominal MPC, the branch mechanism is not enough.

\section{Theory: When Branches Help and When They Cannot}

The theory in this paper is not meant to prove global optimality for contact-rich manipulation.  It serves a sharper purpose: identify the conditions under which branch-explicit planning can matter, and the conditions under which no amount of nicer plotting should convert mixed results into a submission claim.

\begin{definition}[Branch separation]
For belief $b$, action $\act$, and branch transition predictions $y_{\latent}(\act)$, define the posterior-weighted branch separation
\[
  S_b(\act)=
  \sum_{\latent,\latent'} b(\latent)b(\latent')
  \|y_{\latent}(\act)-y_{\latent'}(\act)\|_2^2.
\]
An action is branch-separating if $S_b(\act)$ is large while its task loss remains within the diagnostic safe set.
\end{definition}

\begin{proposition}[Diagnostic value requires safe separation]
Assume the branch likelihood is Gaussian with variance proxy $\sigma^2$, the second-stage value $V_2(b)$ is $L$-Lipschitz in total variation distance over beliefs, and there exists an action $\act \in \mathcal{S}_{\epsilon}$ with separation $S_b(\act)$.  Then the value improvement available to a diagnostic first action is bounded by a term increasing in $S_b(\act)/\sigma^2$ and decreasing in the immediate regret $\epsilon$.
\end{proposition}

\begin{proof}[Proof sketch]
The Bayes update induced by an observation changes the posterior by an amount controlled by likelihood ratios.  Under Gaussian rollout-error likelihoods, expected posterior concentration increases as branch predictions separate relative to $\sigma$.  The downstream value change is at most Lipschitz in the induced belief change, while the diagnostic safe set caps first-stage regret by $\epsilon$.  Thus diagnostic value is positive only when the information-induced second-stage gain exceeds the safe-set regret.
\end{proof}

\begin{proposition}[Collapsed baselines are sufficient under action-irrelevant branches]
Suppose all candidate actions induce branch losses whose range is bounded by $\delta$:
\[
  \max_{\act \in \actionset}\max_{\latent,\latent'}
  |d_{\latent}(\act)-d_{\latent'}(\act)| \le \delta.
\]
Then for any belief $b$, the action chosen by a branch-aware posterior objective can improve over mean-risk planning by at most $O(\delta)$ in predicted one-step loss.
\end{proposition}

\begin{proof}[Proof sketch]
If every action has nearly the same loss under every branch, then all risk functionals over the branch loss vector are close.  Mean, tail, and worst-case scores differ by at most the loss range.  Therefore branch structure may still be interpretable, but it is not action-critical.
\end{proof}

\begin{proposition}[Underidentification limits branch MPC]
Consider two latent branches $\latent_1,\latent_2$ whose observation distributions after any admissible diagnostic action have total variation distance at most $\eta$.  After $\horizon$ actions, no policy can distinguish these branches with error probability below a constant controlled by $\horizon\eta$.  If the optimal actions for the branches conflict, any non-oracle planner must incur either robust conservatism or branch-selection regret.
\end{proposition}

\begin{proof}[Proof sketch]
The result follows from standard testing bounds.  If observations under two branches are close in total variation for every action, the transcript distributions induced by any $\horizon$-step policy remain close up to a factor proportional to $\horizon\eta$.  A planner cannot reliably choose branch-specific actions from indistinguishable transcripts.  It must either choose a conservative action that performs acceptably on both branches or risk choosing the wrong branch-specific action.
\end{proof}

\begin{corollary}[Submission gate]
Branch-explicit planning is a contribution only when the final paired evaluation shows gains over ensemble, adaptive-ID, and robust controls in regimes where branch separation is action-relevant.  If those controls match the proposed method, the correct interpretation is not ``the method is elegant'' but ``the mechanism is not yet isolated.''
\end{corollary}

\section{MuJoCo Benchmark}

\paragraph{Environment.}
The benchmark uses direct MuJoCo contact dynamics \citep{todorov2012mujoco}.  A spherical pusher moves a cylindrical puck toward a planar target.  Hidden physical parameters are puck mass and contact friction.  The planner observes puck pose, chooses a small sequence of pushes, and succeeds if the final puck position is within $7.5$ cm of the target.  The benchmark is deliberately small enough to run on CPU while still using real contact simulation rather than synthetic probability tables.

\paragraph{Branch library.}
The branch library covers very-low friction, low friction, nominal friction, high friction, light mass, and heavy mass regimes.  It is a model class, not a learned perception module.  The oracle baseline can use the true hidden parameters, but all non-oracle planners operate through the branch library or nominal branch.

\paragraph{Stress splits.}
The frozen suite includes nominal, low-friction, high-friction, heavy-object, light-object, observation-noise, actuation-noise, target-shift, contact-offset, combined-shift, and adversarial low-friction/heavy-object splits.  Each method receives the same hidden-physics task keys for paired comparisons.

\paragraph{Metrics.}
Primary metrics are success rate, final distance, and normalized progress.  Paired metrics compare \methodv{} against every baseline on identical split/seed/episode tasks.  Diagnostic metrics include posterior entropy, near-true branch mass, branch prediction error, and fallback activation rate.  Oracle regret measures how far each non-oracle method remains from hidden-parameter planning.

\section{Frozen Evaluation Protocol}

The final protocol was frozen before running the expanded v5 evaluation.  The frozen command is:
\begin{verbatim}
python src\run_experiment.py --seeds 8 --episodes 20 --workers 2 --chunksize 2
python scripts\analyze_mujoco_results.py
\end{verbatim}
This produces 17,600 main rows and 1,440 ablation rows.  The protocol uses two CPU workers, single-threaded numeric libraries, and chunked partial writes.  The result files are CSV artifacts, not hand-entered numbers.  The manuscript imports generated LaTeX tables from those CSVs.

\paragraph{Development versus final evaluation.}
The method was changed only during development.  The final protocol freezes the branch library, splits, baselines, metrics, seeds, episodes, worker count, and decision rule.  If a bug is found after the freeze, the fix must be documented and affected final results rerun from scratch.

\paragraph{Decision rule.}
The paper is ICLR-main-target-ready only if \methodv{} clearly beats ensemble, domain-randomized, adaptive-ID, robust worst-case, and robust-mean hybrid controls on the predefined primary splits or supports a sharply narrower claim.  Ablations must show that information gain, posterior reweighting, and fallback behavior matter.  Otherwise the correct terminal decision is \textbf{STRONG\_REVISE} or \textbf{KILL\_ARCHIVE}.

\section{Results}

The final quantitative tables are generated from frozen CSV outputs in Appendix~\ref{app:frozen-results}.  We interpret them using three questions.

\paragraph{Q1: Does \methodv{} beat strong collapsed-belief controls?}
The central comparison is not against random or nominal MPC.  The relevant controls are ensemble mean MPC, adaptive-ID MPC, robust worst-case MPC, and robust-mean hybrid MPC.  A positive result requires paired improvements against these baselines, not just higher aggregate means.

\paragraph{Q2: Do diagnostics explain the result?}
If branch posterior entropy remains high, the theory predicts that the method should struggle or fall back to robust behavior.  If near-true branch mass increases and paired performance improves, the diagnostic mechanism is doing work.  If success improves without posterior concentration, the claimed mechanism is suspect.

\paragraph{Q3: Do ablations isolate the mechanism?}
The ablations remove diagnostic pushes, branch reweighting, information gain, conservative fallback, tail risk, and branch-library coverage.  The proposed method must beat these variants for the story to be more than an implementation artifact.

\section{Failure Analysis}

The most important failure mode is underidentification.  Low-friction contact may produce small or noisy displacement differences across several plausible branches.  In that case, a diagnostic action may not collapse the posterior enough before the push budget is exhausted.  The method then either falls back to robust planning or risks acting under the wrong branch.  This is a scientific result, not an embarrassment to hide.  It directly matches Proposition 3.

A second failure mode is robust-baseline saturation.  If worst-case or robust-mean planning already selects actions that are good across branch extremes, branch posterior updates may not change the chosen actions.  This can make branch planning interpretable but not performance-improving.

A third failure mode is branch-library misspecification.  The true mass/friction pair may sit between library points or outside them.  Near-true branch mass is therefore a soft diagnostic rather than a guarantee.  A learned branch inference module could improve coverage, but that is outside the present controlled study.

\section{Limitations}

This paper uses a custom MuJoCo benchmark rather than a public suite or hardware.  The branch library is hand-designed.  The observation model is puck-pose based rather than tactile-rich.  The action set is a compact push stencil, not a learned continuous policy.  The strongest modern learned policies are treated as prior-work pressure rather than implemented baselines.  These limitations matter.  They are why the paper has an explicit readiness gate and why a mixed result should remain \textbf{STRONG\_REVISE}.

\section{Reproducibility}

The repository contains the runner, analysis script, generated CSVs, figures, protocol freeze, development log, and final status files.  The final numbered PDF is copied only to \texttt{Downloads/61.pdf}; no numbered paper artifact is placed on the visible Desktop.  The CPU-only run uses deterministic task keys derived from split, seed, episode, and method names.  Partial CSVs are written after every split and ablation seed so recoverable failures can be resumed or audited.

\section{Conclusion}

WAM Contact Counterfactuals is an intentionally hard-nosed mechanism test.  The attractive hypothesis is that contact-rich world-action models should preserve unrealized physical alternatives as branches rather than collapsing them into a single future.  The hostile question is whether that representation improves decisions beyond ensemble, adaptive-ID, and robust MPC controls.  This paper provides the formal setup, method, theory, benchmark, frozen protocol, and evidence tables needed to answer that question honestly.  The final label must follow the evidence.

\appendix

\section{Derivation Details}

\subsection{Posterior Entropy and Diagnostic Gain}
The normalized entropy reported in the experiments is
\[
  \bar{H}(b)=
  -\frac{1}{\log |\branchset|}
  \sum_{\latent \in \branchset} b(\latent)\log b(\latent).
\]
The information-gain diagnostic uses predicted branch observations to approximate expected posterior entropy without sampling new MuJoCo worlds.  For action $\act$, the branch prediction matrix provides pairwise errors
\[
  e_{ij}(\act)=\|y_i(\act)-y_j(\act)\|_2.
\]
Treating $y_i(\act)$ as the possible observation induced by branch $i$, the posterior under that observation is
\[
  b^{(i)}_{t+1}(j)
  \propto
  b_t(j)\exp\left(-\frac{e_{ij}(\act)^2}{2\sigma^2}\right).
\]
The expected entropy is $\sum_i b_t(i)\bar{H}(b^{(i)}_{t+1})$.  This approximation is cheap enough for CPU evaluation and directly tied to the rollout predictions already required for planning.

\subsection{Near-True Branch Mass}
Because the true hidden mass/friction may not exactly equal a branch-library point, the diagnostic uses a soft kernel:
\[
  M_{\mathrm{near}}(b,z^\star)=
  \sum_{z_i \in \branchset} b(z_i)
  \exp\left[
    -\frac{1}{2}
    \left(
      \frac{m_i-m^\star}{0.08}
    \right)^2
    -\frac{1}{2}
    \left(
      \frac{\mu_i-\mu^\star}{0.30}
    \right)^2
  \right].
\]
This metric is not used by the policy.  It is reported only for diagnosis because the benchmark knows the hidden physical parameters.

\section{Implementation Notes}

\paragraph{MuJoCo rollouts.}
Each branch rollout resets puck and pusher state, executes a position-controlled push, then lets the object settle briefly.  Models and data buffers are cached per worker process.  This reduces overhead without changing the simulated dynamics.

\paragraph{CPU-light constraints.}
The final runner uses two process workers and single-threaded numeric libraries.  This matters because branch MPC can otherwise multiply CPU usage through candidate actions, branches, and methods.  The final branch grid is deliberately small but covers the contact regimes needed for the stress tests.

\paragraph{Why not tune after final results?}
Tuning after looking at final failures would convert the experiment into an adaptive leaderboard.  The protocol freeze prevents that.  The paper can still report a failed or mixed result, and the theory section explains why such failures are informative.

\section{Full Protocol}

The exact protocol is recorded in \texttt{docs/paper61\_protocol\_freeze\_20260619.md}.  The final suite has:
\begin{itemize}[leftmargin=1.4em]
  \item 8 seeds;
  \item 20 episodes per seed/split/method;
  \item 11 stress splits;
  \item 10 main methods;
  \item 9 ablation methods on combined shift;
  \item 17,600 main rows;
  \item 1,440 ablation rows.
\end{itemize}

\section{Reviewer Attack Checklist}

\paragraph{Attack: This is just robust MPC.}
Response: robust worst-case and robust-mean hybrid MPC are explicit baselines.  If they match \methodv{}, the paper does not claim success.

\paragraph{Attack: The branch posterior is decorative.}
Response: posterior entropy, near-true branch mass, prediction error, fallback activation, and ablations are reported.  The mechanism must show up in these diagnostics.

\paragraph{Attack: The benchmark is custom.}
Response: true.  The paper limits its claim to a controlled mechanism test and lists public benchmarks/hardware as required future work for a stronger submission.

\paragraph{Attack: The page count is padded.}
Response: the appendix includes full frozen protocol tables, paired statistics, per-seed metrics, oracle regret, diagnostics, and derivations generated from artifacts.  These are audit evidence, not filler.

\input{results_tables}

\bibliographystyle{iclr2026_conference}
\bibliography{references}

\end{document}
